---
title: "Thesis Porfolio 1"
author: "Adam"
date: "4/4/2022"
output: html_document
---

The goal of this markdown is to properly clean data using R.

## Set-up

First, I need to ensure that I have the packages needed.

```{r packages, echo=TRUE, message=FALSE}
library('reshape2')
library('dplyr')
library('Rcpp')
library('readr')
library('tidyverse')
library('psych')
```


### Bringing in the data

```{r bringing in data, echo=TRUE, collapse=TRUE, message=FALSE}
study1 <- read_csv("~/GitHub/Portfolio-1/Portfolio 1 - cleaned (pre).csv")

# Ensuring the dataframe is a tibble.
as_tibble(study1)

#Taking a look at the table.
View(study1)
```



### Checking for bad data

In a unique step, I needed to look at participants to see if any data was suspect, before we could give credit.

Checking for bad actors in the data, before giving credit.

```{r checking the basics}
summary(study1$Durationinseconds)
```

```{r Filtering}
#looking narrowly at people under the 1st quartile (172)

filter(study1, Durationinseconds < 172) %>%
        arrange(Durationinseconds)
```

I found no straight-lining within this subset, which suggests to me that the data was all given in good faith. However, I wanted to look at another subset.

I filtered for students that answered positively on two items that previous research suggest should be opposite, belonging and feeling like an outsider.

```{r}
filter(study1, school_belong > 4, school_outsider > 4)
#40 students fall into this category

filter(study1, class_belong > 4, class_outsider > 4)
#18 students fall into class.

filter(study1, school_belong > 4, school_outsider > 4, class_belong > 4, class_outsider > 4)
#4 students fall into this category. Only one seemed suspicious, but passed along to Shannon to make a decision.

```

Ultimately, we decided to grant credit to all students. That means it's time to move on to cleaning the data.


## Data cleaning

Code to remove identifying information from the dataset. Since I am sharing actual data, the data has already been deidentified in order to comply with the IRB. However, her is the code that was used to do so.


study1 <- subset(study1, select= -c(StartDate, EndDate, Status,	IPAddress, Progress,
                Finished, RecordedDate, ResponseId, RecipientLastName, RecipientFirstName,
                RecipientEmail, ExternalReference, LocationLatitude, LocationLongitude,
                DistributionChannel, UserLanguage))


## Descriptives cleaning

Now it's time to begin creating the dummy codes needed for my analyses.

#### Gender dummy coding

First, I need to look at whether we have non-binary individuals in the dataset.

```{r checking gender demographics}
study1 %>%
        filter(is.na(demog_gender_female), is.na(demog_gender_male)) %>%
        count(demog_gender_trans, demog_gender_twospirit, demog_gender_genderqueer, demog_gender_other, demog_gender_no_answer)

study1 %>%
count(demog_gender_trans, demog_gender_twospirit, demog_gender_genderqueer, demog_gender_other, demog_gender_no_answer)

```

We have 12 people that did not identify as either male or female in some way. Due to this small number, there is not much we can do with the individual gender terms, so I am going to make two variables, gender_male & gender_nonbinary. Women serve as the reference class for the 'binary' gender dummy code.


The code below creates the non-binary variable based on if someone answered yes to any of the demographic non-binary options, and a 0 if they answered to either of the binary answers. Then, the second set of code makes it so that NA answers are coded as non-binary as well.

```{r creating non-binary dummy-code}
study1 <- study1 %>%
        mutate(
        gender_nonbinary=case_when(
        !demog_gender_male & !demog_gender_female & demog_gender_twospirit | demog_gender_genderqueer | demog_gender_other
        ~ 1,
        demog_gender_female | demog_gender_male ~ 0
        ))

# This code will make it so that the NA are coded as 1. For non-binary, those without an answer will be considered non-binary.

study1 <- study1 %>%
        mutate(
        gender_nonbinary=case_when(
                gender_nonbinary == 1 ~ 1,
                gender_nonbinary == 0 ~ 0,
        is.na(gender_nonbinary) ~ 1
))
```

I spent five hours weeks ago trying to get this to work. It went so poorly I chose to make my dummy coded variables in SPSS for my actual thesis. So, I'm pretty pleased to have gotten this to work.

Below, I create the dummy code for the gender binary.

```{r Creating  binary gender dummy code}
study1 <- study1 %>%
        mutate(
        gender_male_dummy=case_when(
                demog_gender_male == 1 ~ 1,
                demog_gender_female | gender_nonbinary ~ 0
                ))
```

        
### Race dummy coding

To create the race dummy codes, I first need to create a multiracial composite so that I can avoid adding multiracial people to individual dummy codes. 

Because the individual race codes have NA in all unfilled spots, I need to replace them with 0's to get the code to work correctly. The code is a little hideous and not consolidated, but I prefer to look and see them individually, especially because I can collapse this code in the chunk.

```{r adding 0 to original code}

study1 <- study1 %>%
        mutate(demog_race_51=case_when(
                demog_race_51 == 1 ~ 1,
                is.na(demog_race_51) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_52=case_when(
                demog_race_52 == 1 ~ 1,
                is.na(demog_race_52) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_59=case_when(
                demog_race_59 == 1 ~ 1,
                is.na(demog_race_59) ~ 0
                                       ))
study1 <- study1 %>%
        mutate(demog_race_21=case_when(
                demog_race_21 == 1 ~ 1,
                is.na(demog_race_21) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_22=case_when(
                demog_race_22 == 1 ~ 1,
                is.na(demog_race_22) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_23=case_when(
                demog_race_23 == 1 ~ 1,
                is.na(demog_race_23) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_29=case_when(
                demog_race_29 == 1 ~ 1,
                is.na(demog_race_29) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_31=case_when(
                demog_race_31 == 1 ~ 1,
                is.na(demog_race_31) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_32=case_when(
                demog_race_32 == 1 ~ 1,
                is.na(demog_race_32) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_33=case_when(
                demog_race_33 == 1 ~ 1,
                is.na(demog_race_33) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_39=case_when(
                demog_race_39 == 1 ~ 1,
                is.na(demog_race_39) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_11=case_when(
                demog_race_11 == 1 ~ 1,
                is.na(demog_race_11) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_12=case_when(
                demog_race_12 == 1 ~ 1,
                is.na(demog_race_12) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_13=case_when(
                demog_race_13 == 1 ~ 1,
                is.na(demog_race_13) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_19=case_when(
                demog_race_19 == 1 ~ 1,
                is.na(demog_race_19) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_41=case_when(
                demog_race_41 == 1 ~ 1,
                is.na(demog_race_41) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_42=case_when(
                demog_race_42 == 1 ~ 1,
                is.na(demog_race_42) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_89=case_when(
                demog_race_89 == 1 ~ 1,
                is.na(demog_race_89) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demog_race_99=case_when(
                demog_race_99 == 1 ~ 1,
                is.na(demog_race_99) ~ 0
                                       ))
```


With that task finished, we can work on making the multiracial composite.


```{r creating a multiracial sum}
study1 <- study1 %>%
        rowwise() %>%
        mutate(
                multiracial_sum=sum(c_across(demog_race_41:demog_race_89
        )))

```

```{r Creating multiracial dummy code}
study1 <- study1 %>%
        mutate(race_multiracial=case_when(
                multiracial_sum > 1 ~ 1,
                multiracial_sum == 1 ~ 0,
                multiracial_sum == 0 ~ 0
        ))

```

Now, I can make the racial composites. Due to sample size constraints I am going to have to make composites for several races. All composites will include their respective races and did not answer more than one race.

#### White composite

Composed of individuals who identified as European/European American, Middle Eastern/Middle Eastern American, Other White; or demog_race_51, demog_race_52, and demog_race_59.

```{r White composite dummy code}
study1 <- study1 %>%
        mutate(
        race_white=case_when(
                demog_race_51 | demog_race_52 | demog_race_59 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_white=case_when(
          race_white == 1 ~ 1,
          race_white == 0 ~ 0,
          is.na(race_white) ~ 0 
          ))

```

#### Black composite

Composed of individuals who identified as African American/Black, African, Caribbean, Other Black; or demog_race_21, demog_race_22, demog_race_23, and demog_race_29.

```{r Black composite dummy code}
study1 <- study1 %>%
        mutate(
        race_black=case_when(
                demog_race_21 | demog_race_22 | demog_race_23 | demog_race_29 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_black=case_when(
          race_black == 1 ~ 1,
          race_black == 0 ~ 0,
          is.na(race_black) ~ 0 
          ))

```

#### Hispanic composite

Composed of individuals who identified as Mexican American/Chicano, Puerto Rican, Central American, Other Hispanic or Latino; or demog_race_31, demog_race_32, demog_race_33, and demog_race_39.

```{r Hispanic composite dummy code}
study1 <- study1 %>%
        mutate(
        race_hispanic=case_when(
                demog_race_31 | demog_race_32 | demog_race_33 | demog_race_39 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_hispanic=case_when(
          race_hispanic == 1 ~ 1,
          race_hispanic == 0 ~ 0,
          is.na(race_hispanic) ~ 0 
          ))

```

#### Asian composite

Composed of individuals who identified as East Asian (e.g., Chinese, Japanese, Korean, Taiwanese), Southeast Asian (e.g., Filipino, Vietnamese, Indonesian), South Asian (e.g., Pakistani, Indian, Nepalese, Sri Lankan), Other Asian; or demog_race_11, demog_race_12, demog_race_13, and demog_race_19.


```{r Asian composite dummy code}
study1 <- study1 %>%
        mutate(
        race_asian=case_when(
                demog_race_11 | demog_race_12 | demog_race_13 | demog_race_19 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_asian=case_when(
          race_asian == 1 ~ 1,
          race_asian == 0 ~ 0,
          is.na(race_asian) ~ 0 
          ))

```

#### Native composite

Composed of individuals who identified as Native Hawaiian or Other Pacific Islander, American Indian or Alaskan Native; or demog_race_41 and demog_race_42.



```{r Native composite dummy code}
study1 <- study1 %>%
        mutate(
        race_native=case_when(
                demog_race_41 | demog_race_42 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_native=case_when(
          race_native == 1 ~ 1,
          race_native == 0 ~ 0,
          is.na(race_native) ~ 0 
          ))

```

With these in hand, I should make sure they are all necessary.

```{r freq race, collapse=TRUE}
table(study1$race_white)
table(study1$race_black)
table(study1$race_asian)
table(study1$race_hispanic)
table(study1$race_native)
table(study1$race_multiracial)
```

We have participants of every racial demographic.



## Variables of Interest

Next, we need to create the composite variables for the variables of interest.


### Institutional belonging

There are four variables that make up institutional belonging, "school_belong", "school_outsider", "school_comfort", "school_fit_in".

One of the four variables needs to be reverse scored, school_outsider.

```{r reverse scoring school outsider}
study1 <- study1 %>%
        mutate(school_outsiderR= 8- school_outsider
                              )
```

Testing the reliability of the measure.

```{r institutional belonging reliability}
alpha(study1[,c("school_belong", "school_outsiderR", "school_comfort", "school_fit_in")])

```

Reliability is .88.

Creating the actual composite.

```{r institutional belonging composite}

study1$school_belong_comp4 <- rowMeans(study1[,c("school_belong", "school_outsiderR", "school_comfort", "school_fit_in")], 
                              na.rm=TRUE)
```


> This explains how the above code works. I took it from someone else online, and it is here for my own edification (https://rforhr.com/compositevariable.html).

Type the name of the rowMeans function. 
As the first argument, type the name of the data frame object to which the items belong (study1). 
Following the data frame name, type in brackets ([ ]), and within the brackets, type a comma (,) followed by the c (combine) function; by placing a comma in front of the c function, we are indicating that we will be referencing the names of columns (i.e., variables); within the c function, list the name of each item in quotation marks (" "), separated by commas (,). 
 Finally, as the second argument in the rowMeans function, type the na.rm=TRUE argument, which will tell the function to compute the mean for each case that has at least one score for the specified items; in other words, this function allows for the row means to be computed even if there are missing data.

I know that I have no missing responses, but it's good to have that there for future use. And, it doesn't change anything because there's no one missing.

### Class belonging

Same as institutional belonging, there are the same four questions (with the prefix class_ this time) and I need to reverse score "class_outsider".

```{r reverse scoring class outsider}
study1 <- study1 %>%
        mutate(class_outsiderR= 8- class_outsider
        )

```

```{r class belonging reliability}
alpha(study1[,c("class_belong", "class_outsiderR", "class_comfort", "class_fit_in")])
```

Reliability is .91.

```{r Class belonging composite}
study1$class_belong_comp4 <- rowMeans(study1[,c("class_belong", "class_outsiderR", "class_comfort", "class_fit_in")], 
                              na.rm=TRUE)
```



### Class disclosure

Similar to above, I need to make a composite of two variables: "cdiscl_private" and "cdiscl_self_talk".

```{r class disclosure reliability}
alpha(study1[,c("cdiscl_private", "cdiscl_self_talk")])
```

Reliability is .89.

```{r Class disclosure composite}
study1$class_disclose_comp2 <- rowMeans(study1[,c("cdiscl_private", "cdiscl_self_talk")], 
                              na.rm=TRUE)
```



### Class prompted disclosure

Composite made of two variables, "cdiscl_prompt_prof" and "cdiscl_prompt_ecour".

```{r class prompted disclosure reliability}
alpha(study1[,c("cdiscl_prompt_prof", "cdiscl_prompt_ecour")])
```

Reliability is .83.

```{r Class prompted disclosure composite}
study1$cdiscl_prompt_comp2 <- rowMeans(study1[,c("cdiscl_prompt_prof", "cdiscl_prompt_ecour")], 
                              na.rm=TRUE)
```


### Group disclosure.

Reports on how much participants disclosed to their group. Composite made of two variables; "gdiscl_private" and "gdiscl_self_talk". 

This is the first composite to have missing variables, and it has a lot of them, because a large portion of our sample did not use groups. This will be discussed in portfolio 2.

```{r group disclosure reliability}
alpha(study1[,c("gdiscl_private", "gdiscl_self_talk")])
```

Reliability is .90.

```{r Group disclosure composite}
study1$group_disclose_comp2 <- rowMeans(study1[,c("gdiscl_private", "gdiscl_self_talk")], 
                              na.rm=TRUE)
```


### Group prompted disclosure 

Composite made of two variables, "gdiscl_prompt_prof" and "gdiscl_prompt_encour". Also has missing variables.

```{r group prompted disclosure reliability}
alpha(study1[,c("gdiscl_prompt_prof", "gdiscl_prompt_encour")])
```

Reliability is .79.

```{r Group prompted disclosure composite}
study1$gdiscl_prompt_comp2 <- rowMeans(study1[,c("gdiscl_prompt_prof", "gdiscl_prompt_encour")], 
                              na.rm=TRUE)
```

With that, we've finished creating the composite variables! There is no more data cleaning that needs to be done, so this portfolio is finished.


## Saving the data!

```{r saving data}
write.csv(study1, file = "Study1_data_R_cleaned.csv")
```


### Wrapping up

I don't like data cleaning in R. It's clunky, weird, and makes it a lot harder than it needs to be. I've found out that most of my committee cleans their data in SPSS then moves it to R (or Mplus in Keita's case). I think I'm going to do the same, but I'm glad I know how to do this now.